A Large and Variable Leading Tail of Helium in a Sub-Saturn Undergoing Runaway Inflation

### 2023-06-27 paper/paper1/ms.tex
Up to page 10

We masked spectral regions predicted to have significant telluric lines with a template generated by \texttt{telfit} \citep{2014AJ....148...53G}.  We shifted the spectral coordinates to their common barycentric-corrected reference frame \citep{2014PASP..126..838W} as implemented in \texttt{astropy} \citep{2013A&A...558A..33A,2018AJ....156..123A,2022ApJ...935..167A}.  We flattened the spectra to two pre-selected continuum indices highlighted as vertical blue bands in Figure \ref{fig:HPFheliumOverview}.  The sky-subtraction, telluric masking, barycentric correction, flattening, and all of our other standard pre-processing steps were carried out in the open-source Python interface \texttt{muler} \citep{2022JOSS....7.4302G}.

Figure \ref{fig:HPFheliumOverview} shows a zoom-in on the He 10830 region of interest, with all 152 individual exposures overlaid.  We see variability in the Helium line of up to 10\%, much greater than the $<1\%$ pixel-to-pixel variation.  The feature width spans about 3~\AA.

@@

\begin{figure}
    \includegraphics[width=\linewidth]{figures/phase_2D_diagram_resid.png}
    \caption{Absorption depth phase scan over the entire HAT-P-67b orbit from 41 visits with HPF.  }
    \label{fig:HPFscanResid}
\end{figure}

@@

\subsection{Size and geometry of outflow}
Helium absorption occurs predominantly before the planet is in-transit, indicating a \emph{leading} tail escaping the planet.  The leading tail is detected 0.1 AU away from the planet, or 130 planetary radii, well outside of the Roche Lobe at $<2.6$ planetary radii.  We are detecting Roche lobe overflow, escaping material that is entirely unbound from the planet.  The bulk of the detectable unbound gas may be understood as following its own Keplerian trajectory at a slightly shorter period orbit, pulling away from the planet parallel to the orbital path.  Such a scenario may arise from preferential emission on the planet dayside.

@@

>>>>

We masked spectral regions predicted to have significant telluric lines with a template generated by \texttt{telfit} \citep{2014AJ....148...53G} then shifted the spectral coordinates to their common barycentric-corrected reference frame \citep{2014PASP..126..838W}, as implemented in \texttt{astropy} \citep{2013A&A...558A..33A,2018AJ....156..123A,2022ApJ...935..167A}.  The spectral continuum was flattened and normalized to two pre-selected continuum indices highlighted as vertical blue bands in Figure \ref{fig:HPFheliumOverview}.  The sky-subtraction, telluric masking, barycentric correction, flattening, and all of our other standard pre-processing steps were carried out in the open-source Python interface \texttt{muler} \citep{2022JOSS....7.4302G}.

Figure \ref{fig:HPFheliumOverview} shows a zoom-in on the He 10830 \AA~ region of interest, with all 152 individual exposures overlaid.  We see variability in the Helium line of up to 10\%, much greater than the $<1\%$ pixel-to-pixel variation.  The feature width spans about 3~\AA.

@@

\begin{figure}
    \includegraphics[width=\linewidth]{figures/phase_2D_diagram_resid.png}
    \caption{Absorption depth phase scan over the entire HAT-P-67b orbit from 41 visits with HPF.  The planet orbital rest frame is shown as the black sine wave. Gaps in data at 10834-10836 \AA arise from telluric masking.  Unmasked telluric lines are perceptible at 10825 \AA.}
    \label{fig:HPFscanResid}
\end{figure}

@@

\subsection{Size and geometry of outflow}
Helium absorption occurs predominantly before the planet is in-transit, indicating a \emph{leading} tail escaping the planet.  The leading tail is detected 0.1 AU away from the planet, or 130 planetary radii, well outside of the Roche lobe at $<2.6$ planetary radii.  We are detecting Roche lobe overflow, escaping material that is entirely unbound from the planet.  The bulk of the detectable unbound gas may be understood as following its own Keplerian trajectory at a slightly shorter period orbit, pulling away from the planet parallel to the orbital path.  Such a scenario may arise from preferential emission on the planet dayside.

@@

The leading tail geometry seen in in HAT-P-67~b contrasts with the morphology of the HAT-P-32~b system, which exhibits giant symmetric tidal tails with comparable absorption depth preceding and trailing the planet \citep{doi:10.1126/sciadv.adf8736}.  These systems represent two of the most extended features associated with exoplanets, and we compare and contrast them in the Section \ref{secDiscuss}.
